---
title: "Figure 2"
output: html_document
date: "2025-07-28"
---

```{r setup}
base_dir<-"~/Desktop/mount2/chidalgo/pdzext/"
knitr::opts_knit$set(root.dir = base_dir)
setwd(base_dir)
```


#R packages
```{r}
library(data.table)
library(ggplot2)
library(ggpubr)
library("readxl")
library(ggrepel)
library(scales)
library(ggtext)
```


#tables used
```{r}
mutations_distances_annotation_median_rSASA_Folding_merged<-fread("Data_pdzext/energies_pdzext/thirdOrder_mutations_distances_annotation_median_rSASA_Folding_merged.txt")
mutations_distances_annotation_median_rSASA_Binding_merged<-fread("Data_pdzext/energies_pdzext/thirdOrder_mutations_distances_annotation_median_rSASA_Binding_merged.txt")
signif_muts_bysite_distances_annotation_median_rSASA_Binding_merged<-fread("Data_pdzext/energies_pdzext/thirdOrder_signif_muts_bysite_distances_annotation_median_rSASA_Binding_merged.txt")
signif_muts_bysite_distances_annotation_median_rSASA_Folding_merged<-fread("Data_pdzext/energies_pdzext/thirdOrder_signif_muts_bysite_distances_annotation_median_rSASA_Folding_merged.txt")
```

```{r}
raw_list_Folding<-fread("Data_pdzext/energies_pdzext/thirdOrder_merge_table_Folding.txt")[
  id %in% c("WT", "C67A", "C68A", "C67A_C68A"), 
  .(id, total_mean_kcal_mol, total_std_kcal_mol)
]

raw_list_Binding<-fread("Data_pdzext/energies_pdzext/thirdOrder_merge_table_Binding.txt")[
  id %in% c("WT", "C67A", "C68A", "C67A_C68A"), 
  .(id, total_mean_kcal_mol, total_std_kcal_mol)
]
```

```{r}
#load tables with literature energy measurements
PDZ_literature_free_energies<-fread("Data_pdzext/literature_validations_pdzext/PDZ_literature_free_energies.txt")

saotome2021_invitro<-data.table(Dataset=c("Saotome_2021","Saotome_2021","Saotome_2021","Saotome_2021","Saotome_2021","Saotome_2021","Saotome_2021"),id=c("V328A","F340A","L342A","I389A","V315A","L349A","V365A"),Tm=c(68.97,72.98,70.97,60.99,66.21,69.74,70.01),dHvanHoff=c(262.47,350.08,339.13,237.23,224.99,241.75,179.84))

PDZ_literature_free_energies<-rbind(PDZ_literature_free_energies,saotome2021_invitro,fill=TRUE)

single_dG_Salinas_2018_pdz3_long<-fread("Data_pdzext/literature_validations_pdzext/single_dG_Salinas_2018_pdz3_long.txt")

ddPCApaper_folding<-fread("Data_pdzext/literature_validations_pdzext/DDG_folding_ddPCAPaper_newMochi.txt")
ddPCApaper_folding[,Pos:=Pos.y-311]
ddPCApaper_folding[,id:=paste0(WT_AA,Pos,Mut)]
ddPCApaper_binding<-fread("Data_pdzext/literature_validations_pdzext/DDG_binding_ddPCAPaper_newMochi.txt")
ddPCApaper_binding[,Pos:=Pos.y-311]
ddPCApaper_binding[,id:=paste0(WT_AA,Pos,Mut)]
```

#plots Figure 2c and Extended Data 2e: correlation with literature energies
```{r}
#my data
table_first_Folding<-mutations_distances_annotation_median_rSASA_Folding_merged[dataset=="1st_order_ddg",c("wt_Pos","mutant","total_mean_kcal_mol")]
table_first_Folding<-table_first_Folding[,id:=paste0(wt_Pos,mutant)]

table_first_Binding<-mutations_distances_annotation_median_rSASA_Binding_merged[dataset=="1st_order_ddg",c("wt_Pos","mutant","total_mean_kcal_mol")]
table_first_Binding<-table_first_Binding[,id:=paste0(wt_Pos,mutant)]
```

```{r}
#merge mine with literature data
ddg_Folding_first_literature<-merge(table_first_Folding,PDZ_literature_free_energies,by="id")
ddg_Binding_first_literature<-merge(table_first_Binding,PDZ_literature_free_energies,by="id")
```

```{r}
plot_configs <- list(
  list(data = ddg_Folding_first_literature, dataset = "Saotome_2021", x = "dHvanHoff", assay = "Folding", title = "Folding ddg vs in vitro Saotome2021 van't Hoff enthalpy (ΔH)", file = "Data_pdzext/output_pdzext/Fig_2c_saotome.pdf"),
  list(data = ddg_Folding_first_literature, dataset = "Calosci_2008", x = "f_ddg", assay = "Folding", title = "Folding ddg vs in vitro Calosci_2008 f_ddg", file = "Data_pdzext/output_pdzext/Fig_2c_calosci.pdf"),
  list(data = ddg_Folding_first_literature, dataset = "Chi_2008",     x = "f_ddg", assay = "Folding", title = "Folding ddg vs in vitro Chi_2008 f_ddg", file = "Data_pdzext/output_pdzext/Fig_2c_chi_folding.pdf"),
  list(data = ddg_Binding_first_literature, dataset = "Laursen_2020", x = "log10(Kd)", assay = "Binding", title = "Binding ddg vs in vitro Laursen_2020 log10(Kd)", file = "Data_pdzext/output_pdzext/Fig_2c_Laursen.pdf"),
  list(data = ddg_Binding_first_literature, dataset = "Chi_2008",     x = "log10(Kd)", assay = "Binding", title = "Binding ddg vs in vitro Chi_2008 log10(Kd)", file = "Data_pdzext/output_pdzext/Fig_2c_Chi_binding.pdf")
)

# Calculate global y-axis limits across all datasets
all_values <- c(
  ddg_Folding_first_literature$total_mean_kcal_mol,
  ddg_Binding_first_literature$total_mean_kcal_mol
)
global_ylim <- range(all_values, na.rm = TRUE)
y_margin <- 0.1 * diff(global_ylim)
global_ylim_extended <- c(global_ylim[1] - y_margin, global_ylim[2] + y_margin)

for (cfg in plot_configs) {
  data <- cfg$data[Dataset == cfg$dataset]

  p <- ggplot(data, aes_string(x = cfg$x, y = "total_mean_kcal_mol")) +
    geom_smooth(method = "lm", se = FALSE, color = "lightblue", linewidth = 3) +
    geom_point(size = 3) +
    stat_cor(size = 5) +
    coord_cartesian(ylim = global_ylim_extended) +
    theme_classic() +
    theme(aspect.ratio = 1) +
    ggtitle(cfg$title)
  
  print(p)
  ggsave(cfg$file, plot = p, width = 4, height = 4)
}
```

```{r}
#correlation with Salinas et al.(2018)
mutations_distances_annotation_median_rSASA_Binding_merged_filtered <- mutations_distances_annotation_median_rSASA_Binding_merged[dataset == "1st_order_ddg", .(id, total_mean_kcal_mol)]
merged_dt <- merge(single_dG_Salinas_2018_pdz3_long, mutations_distances_annotation_median_rSASA_Binding_merged_filtered[, .(id, total_mean_kcal_mol)], by = "id", all = FALSE)

  p <- ggplot(merged_dt, aes(x = Salinas_2018_single_dG, y = total_mean_kcal_mol)) +
    geom_smooth(method = "lm", se = FALSE, color = "lightblue", linewidth = 3) +
    geom_point(size = 3) +
    stat_cor(size = 5) +
    coord_cartesian(ylim = global_ylim_extended) +
    theme_classic() +
    theme(aspect.ratio = 1) +
    ggtitle("Salinas2018")
  
  print(p)
  ggsave("Data_pdzext/output_pdzext/Fig_2c_salinas.pdf", plot = p, width = 4, height = 4)
```

```{r}
#correlation with Faure et al. (2022) folding
mutations_distances_annotation_median_rSASA_Folding_merged_filtered <- mutations_distances_annotation_median_rSASA_Folding_merged[dataset == "calculated_ddg_NC", .(id, total_mean_kcal_mol)]
merged_dt_folding <- merge(ddPCApaper_folding, mutations_distances_annotation_median_rSASA_Folding_merged_filtered[, .(id, total_mean_kcal_mol)], by = "id", all = FALSE)

  p <- ggplot(merged_dt_folding, aes(x = PDZ_mean_kcal.mol, y = total_mean_kcal_mol)) +
    geom_smooth(method = "lm", se = FALSE, color = "lightblue", linewidth = 3) +
   geom_point(size = 3, alpha=0.5) +
    stat_cor(size = 5) +
    theme_classic() +
    theme(aspect.ratio = 1) +
    ggtitle("ddPCA_folding")
  
  print(p)
  ggsave("Data_pdzext/output_pdzext/Suppl_Fig_2e_folding.pdf", plot = p, width = 4, height = 4)
```

```{r}
#correlation with Faure et al. (2022) binding
mutations_distances_annotation_median_rSASA_Binding_merged_filtered <- mutations_distances_annotation_median_rSASA_Binding_merged[dataset == "calculated_ddg_NC", .(id, total_mean_kcal_mol)]
merged_dt_binding <- merge(ddPCApaper_binding, mutations_distances_annotation_median_rSASA_Binding_merged_filtered[, .(id, total_mean_kcal_mol)], by = "id", all = FALSE)

  p <- ggplot(merged_dt_binding, aes(x = PDZ_mean_kcal.mol, y = total_mean_kcal_mol)) +
    geom_smooth(method = "lm", se = FALSE, color = "lightblue", linewidth = 3) +
    geom_point(size = 3, alpha=0.5) +
    stat_cor(size = 5) +
    theme_classic() +
    theme(aspect.ratio = 1) +
    ggtitle("ddPCA_binding")
  
  print(p)
  ggsave("Data_pdzext/output_pdzext/Suppl_Fig_2e_binding.pdf", plot = p, width = 4, height = 4)
```

#plot Figure 2e: heat maps of energies
```{r}
assays<-c("Folding","Binding")
dataset_values<-c("1st_order_ddg", "dddg_N","dddg_C","ddddg_NC")

aa_color_map <- c(
  "T" = "<span style='color:orange'>T</span>",
  "S" = "<span style='color:orange'>S</span>",
  "N" = "<span style='color:orange'>N</span>",
  "Q" = "<span style='color:orange'>Q</span>",
  "D" = "<span style='color:red'>D</span>",
  "E" = "<span style='color:red'>E</span>",
  "K" = "<span style='color:blue'>K</span>",
  "R" = "<span style='color:blue'>R</span>",
  "H" = "<span style='color:blue'>H</span>",
  "G" = "<span style='color:violet'>G</span>",
  "P" = "<span style='color:violet'>P</span>",
  "C" = "<span style='color:violet'>C</span>",
  "M" = "<span style='color:pink'>M</span>",
  "A" = "<span style='color:pink'>A</span>",
  "L" = "<span style='color:pink'>L</span>",
  "I" = "<span style='color:pink'>I</span>",
  "V" = "<span style='color:pink'>V</span>",
  "F" = "<span style='color:pink'>F</span>",
  "Y" = "<span style='color:pink'>Y</span>",
  "W" = "<span style='color:pink'>W</span>"
)


assay_scale_limits <- list()#Create a list to store global min and max per assay
for (assay in assays) {
  name <- paste0("mutations_distances_annotation_median_rSASA_",assay,"_merged")
    dt_all <- get(name)
    dt_filtered<-dt_all[dataset %in% dataset_values]
    assay_scale_limits[[assay]] <- range(dt_filtered$total_mean_kcal_mol, na.rm = TRUE)#Compute min and max across all datasets for this assay

for (dataset_value in dataset_values) {

subset_dt <- dt_filtered[dataset == dataset_value]
dt_new <- copy(subset_dt)

min_val <- assay_scale_limits[[assay]][1]
max_val <- assay_scale_limits[[assay]][2]

# Extract unique Pos and wt combinations
wt_rows <- unique(dt_new[, .(Pos, wt, dataset)])

#  Create the new rows with total_mean_kcal_mol = NA
wt_rows[, `:=`(
  mutant = wt,
  total_mean_kcal_mol = NA
)]
# Add colored label
    wt_rows[, mutant_colored := aa_color_map[wt]]

# Bind to the original table
dt_final <- rbind(dt_new, wt_rows, fill = TRUE)
    
dt_final$mutant_colored<-factor(dt_final$mutant_colored,levels=c(
  "<span style='color:orange'>S</span>",
  "<span style='color:orange'>T</span>",
  "<span style='color:orange'>N</span>",
  "<span style='color:orange'>Q</span>",
  "<span style='color:red'>D</span>",
  "<span style='color:red'>E</span>",
  "<span style='color:blue'>K</span>",
  "<span style='color:blue'>R</span>",
  "<span style='color:blue'>H</span>",
  "<span style='color:violet'>G</span>",
  "<span style='color:violet'>P</span>",
  "<span style='color:violet'>C</span>",
  "<span style='color:pink'>M</span>",
  "<span style='color:pink'>A</span>",
  "<span style='color:pink'>L</span>",
  "<span style='color:pink'>I</span>",
  "<span style='color:pink'>V</span>",
  "<span style='color:pink'>F</span>",
  "<span style='color:pink'>Y</span>",
  "<span style='color:pink'>W</span>"))    


rSASA_class_df <- mutations_distances_annotation_median_rSASA_Folding_merged[dataset == "1st_order_ddg", .(Pos, wt, class = class_rSASA_Folding)]#to have the coloured x axis
rSASA_class_colors <- c(
  "core" = "#94da55",
  "surface" = "#17A2B8"
)
rSASA_class_df[, wt_colored := paste0(
  "<span style='color:", rSASA_class_colors[class], "'>", wt, "</span>"
)]


  pos_df<-subset_dt[,c("Pos","wt_Pos","wt")]
  pos_df<-pos_df[!duplicated(pos_df),]
  pos_df<-pos_df[order(pos_df$Pos),]
  
  pos_df <- merge(pos_df, rSASA_class_df[, .(Pos, wt_colored)], by = "Pos", all.x = TRUE)
 
  
  p<-ggplot(dt_final,aes(x=Pos,y=mutant_colored))+
     geom_tile(aes(fill = total_mean_kcal_mol))+
   scale_fill_gradientn(colours = c("#898ac0","#f5f5f5","#e85d47"),na.value = "#f5f5f5", values = rescale(c(-max_val,0,max_val)),guide = "colorbar", limits=c(-max_val,max_val))+
   geom_text(
    data = dt_final[is.na(total_mean_kcal_mol)],
    aes(label = "-"),
    color = "black", size = 5, na.rm = TRUE
  ) +
     scale_x_discrete(limits=pos_df$Pos,
                  labels=pos_df$wt_colored)+
   # geom_point(
  #  data = dt_final[class %in% c("stabilizing", "destabilizing")],
   # aes(x = Pos, y = mutant_colored),
  #  shape = 18, color = "black", size = 0.5
  #) +
     theme_classic()+
  theme(
  legend.key.size = unit(0.5, "cm"),  # Adjust the legend key size
  legend.text = element_text(size = 8),
  axis.text.x = element_markdown(vjust = 0.5, hjust = 1),
  axis.text.y = element_markdown(angle = 90,vjust = 0.5, hjust = 1),
  legend.title = element_text(size = 0))+
       coord_fixed(ratio = 1)+
  ggtitle(paste0(assay," energy term ", dataset_value))
  
print(p)
ggsave(paste0("Data_pdzext/output_pdzext/Fig_2e_",dataset_value,"_",assay,".pdf"), width=15, height=5)
}
}
```


#plot figure 2d box plots of energy terms(mutations)
```{r}
dataset_values<-c("ddddg_NC", "dddg_C", "dddg_N","1st_order_ddg")

for (assay in assays) {
  name <- paste0("mutations_distances_annotation_median_rSASA_",assay,"_merged")
    dt <- get(name)

  subset_dt <- dt[dataset %in% dataset_values]
  subset_dt[, dataset := factor(dataset, levels = c("ddddg_NC", "dddg_C", "dddg_N","1st_order_ddg"))]
  
  name2 <- paste0("raw_list_",assay)
    dt2 <- get(name2)

p<-ggplot(subset_dt, aes(x=total_mean_kcal_mol, y=dataset))+

geom_jitter(aes(color=class), size = 1, alpha = 0.3,width = 0.005,height = 0.15)+
  scale_color_manual(
    name = "Point Color",
    values = c("destabilizing" = "#e85d47", "stabilizing" = "#898ac0", "no change" = "grey")
  ) +
 geom_boxplot(width = 0.2, outlier.shape = NA, fill = NA, color = "black", size = 0.2) +
  scale_x_continuous(limits = c(-2, 3.1))+
    geom_vline(# Add one vertical line per dataset
      data = dt2,
      aes(xintercept = total_mean_kcal_mol, color=id),
      
      linetype = "dashed",
      linewidth = 0.2
    ) +
   geom_text(
      data = dt2,
      aes(x = total_mean_kcal_mol, y=1,label = id, color=id),
      hjust = -0.1,  # position label slightly to the right
      size = 3
    ) +
  theme_classic() +
   theme(aspect.ratio=1)+
  ggtitle(paste0(assay, "density"))

  print(p)
ggsave(paste0("Data_pdzext/output_pdzext/Fig_2d_",assay,".pdf"), width=7, height=7)
}
```


#plot Extended Data Fig. 2a: violin plots core/surface/binding
```{r}
custom_colors <- c("surface" = "#17A2B8", "core" = "#94da55", "binding_interface"="#ff7f7f")

for (assay in assays) {
  name <- paste0("mutations_distances_annotation_median_rSASA_",assay,"_merged")
    dt <- get(name)
    subset_dt <- dt[dataset== "1st_order_ddg"]
    class_column<-paste0("class_rSASA_",assay)
    subset_dt[[class_column]] <- factor(
  subset_dt[[class_column]],
  levels = c("core", "surface", "binding_interface")
)
     # Set comparisons for stat_compare_means
  if (assay == "Folding") {
    comparisons_list <- list(c("core", "surface"))
  } else {
    comparisons_list <- list(
      c("core", "binding_interface"),
      c("surface", "binding_interface")
    )
  }
    y_max <- max(subset_dt$total_mean_kcal_mol, na.rm = TRUE)
y_lim <- y_max * 1.5  # or add a fixed buffer if you prefer: y_max + 1

p <- ggplot(subset_dt, aes(x = .data[[class_column]], y = total_mean_kcal_mol)) +
  geom_jitter(aes(color = .data[[class_column]]), width = 0.2, alpha = 0.7, size = 0.1) +
  geom_violin(aes(fill = .data[[class_column]]), trim = TRUE, alpha = 0.3, size = 0.1) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  stat_compare_means(method = "t.test", 
                           comparisons = comparisons_list,
                           label = "p.signif",
                           size = 10)+
 scale_y_continuous(limits = c(min(subset_dt$total_mean_kcal_mol, na.rm = TRUE), y_lim)) +
  theme_classic() +
   theme(aspect.ratio=1)+
  ggtitle(assay)

  print(p)
ggsave(paste0("Data_pdzext/output_pdzext/Suppl_Fig_2a_",assay,".pdf"), width=7, height=7)
}
```



#plot Extended figure 2b: correlation claculated_ddg to see changes in sign
```{r}
library(ggrepel)
 dataset_all<-c("1st_order_ddg", "calculated_ddg_N", "calculated_ddg_C", "calculated_ddg_NC")

for (assay in assays) {
  name <- paste0("signif_muts_bysite_distances_annotation_median_rSASA_",assay,"_merged")
    dt <- get(name)
 dt_base <- dt[dataset == "1st_order_ddg",
                .(Pos, median_energy)]
  setnames(dt_base, "median_energy", "energy_base")

for (comp_ds in dataset_all[-1]) {
    dt_comp <- dt[dataset == comp_ds,
                  .(Pos, median_energy)]
    setnames(dt_comp, "median_energy", "energy_comp")
    dt_merged <- merge(dt_base, dt_comp, by = "Pos")#Merge on Pos (inner join)

    # Skip if no overlap
    if (nrow(dt_merged) == 0L) next

    # Build the plot
y_lims <- if (assay == "Folding") c(-0.5, 2.5) else c(-0.25, 1.25)
x_lims <- if (assay == "Folding") c(-0.5, 2.5) else c(-0.25, 1.25)
    p <- ggplot(dt_merged,
                aes(y = energy_base, x = energy_comp)) +
      geom_point(color = "black", alpha = 0.65) +
     geom_abline(slope = 1, intercept = 0, colour = "blue", linetype = "dashed")+
      geom_hline(yintercept = 0, linetype = "dotted", color = "gray50") +
  geom_vline(xintercept = 0, linetype = "dotted", color = "gray50") +
     stat_cor(method = "pearson",
           label.y = 0.75, label.x = 0.25, size = 6) +
      
        # Label only the points with opposite signs
  geom_text_repel(
    data = dt_merged[(energy_base > 0.05 & energy_comp < -0.05) |
                 (energy_base < -0.05 & energy_comp > 0.05)],
    aes(label = Pos),
    size = 3, color = "red", max.overlaps = Inf
  ) +
       scale_y_continuous(limits = y_lims) + 
       scale_x_continuous(limits = x_lims) + 
      
      theme_classic() +
      theme(aspect.ratio = 1,
            axis.text   = element_text(size = 12),
            axis.title  = element_text(size = 14),
            plot.title  = element_text(size = 15, face = "bold")) +
      labs(
        y     = "1st_order_ddg energy (kcal/mol)",
        x     = paste0(comp_ds, " energy (kcal/mol)"),
        title = paste(assay, "— 1st_order_ddg vs", comp_ds)
      )
    print(p)
    ggsave(paste0("Data_pdzext/output_pdzext/Suppl_Fig_2b_",comp_ds,"_",assay,".pdf"), width=7, height=7)
}
}
```

#plot Figure 2f: paired dot plots of mutations for 337 comparing N-PDZ-C&N-PDZ and 368 comparing N-PDZ-C&PDZ
```{r}
#337
  name <- paste0("mutations_distances_annotation_median_rSASA_Folding_merged")
    dt <- get(name)
    subset_dt<- dt[Pos==337&dataset%in%c("1st_order_ddg","calculated_ddg_C"),]
    
     # Ensure both datasets are present per id
  valid_ids <- subset_dt[, .N, by = id][N == 2, id]
  subset_dt <- subset_dt[id %in% valid_ids]
  
  # Perform paired t-test directly
  value_ordered <- subset_dt[order(id, dataset)]
  t_test_result <- t.test(
    x = value_ordered[dataset == "1st_order_ddg", total_mean_kcal_mol],
    y = value_ordered[dataset == "calculated_ddg_C", total_mean_kcal_mol],
    paired = TRUE
  )
  print(t_test_result)
  
  p_value_label <- paste0("Paired t-test p = ", signif(t_test_result$p.value, 3))
  
  subset_dt[, color_condition := fifelse(dataset == "1st_order_ddg" & fdr_stab_destab < 0.05 & total_mean_kcal_mol > 0, "positive",
                                 fifelse(dataset == "calculated_ddg_C" & fdr_stab_destab < 0.05 & total_mean_kcal_mol < 0, "negative", "other"))]
    
 
p <- ggplot(subset_dt, aes(x = dataset, y = total_mean_kcal_mol)) +
 geom_point(aes(color = color_condition), width = 0.2, alpha = 0.7, size = 5) +
    scale_color_manual(values = c("positive" = "#e85d47", "negative" = "#898ac0", "other" = "grey")) +
   geom_boxplot(width = 0.2, outlier.shape = NA, fill = NA, color = "black", size = 0.5) +
  geom_line(aes(group = id), color = "black", alpha = 0.3) +
   scale_y_continuous(limits = c(-0.7,2.1)) + 
   labs(title = assay, subtitle = p_value_label)+
  theme_classic() +
   theme(aspect.ratio=1)+
  ggtitle(assay)

  print(p)
    ggsave(paste0("Data_pdzext/output_pdzext/Fig_2f_337.pdf"), width=7, height=7)
```

```{r}
#368
  name <- paste0("mutations_distances_annotation_median_rSASA_Folding_merged")
    dt <- get(name)
    subset_dt<- dt[Pos==368&dataset%in%c("1st_order_ddg","calculated_ddg_NC"),]
    
     # Ensure both datasets are present per id
  valid_ids <- subset_dt[, .N, by = id][N == 2, id]
  subset_dt <- subset_dt[id %in% valid_ids]
  
  # Perform paired t-test directly
  value_ordered <- subset_dt[order(id, dataset)]
  t_test_result <- t.test(
    x = value_ordered[dataset == "1st_order_ddg", total_mean_kcal_mol],
    y = value_ordered[dataset == "calculated_ddg_NC", total_mean_kcal_mol],
    paired = TRUE
  )
  print(t_test_result)
  
  p_value_label <- paste0("Paired t-test p = ", signif(t_test_result$p.value, 3))
  
  subset_dt[, color_condition := fifelse(dataset == "1st_order_ddg" & fdr_stab_destab < 0.05 & total_mean_kcal_mol < 0, "negative",
                                 fifelse(dataset == "calculated_ddg_NC" & fdr_stab_destab < 0.05 & total_mean_kcal_mol > 0, "positive", "other"))]
    
p <- ggplot(subset_dt, aes(x = dataset, y = total_mean_kcal_mol)) +
  geom_point(aes(color = color_condition), alpha = 0.7, size = 5) +
    scale_color_manual(values = c("positive" = "#e85d47", "negative" = "#898ac0", "other" = "grey")) +
   geom_boxplot(width = 0.2, outlier.shape = NA, fill = NA, color = "black", size = 0.5) +
  geom_line(aes(group = id), color = "black", alpha = 0.3) +
    scale_y_continuous(limits = c(-0.7,2.1)) + 
   labs(title = assay, subtitle = p_value_label)+
  theme_classic() +
   theme(aspect.ratio=1)+
  ggtitle(assay)

  print(p)
    ggsave(paste0("Data_pdzext/output_pdzext/Fig_2f_368.pdf"), width=7, height=7)
```

