---
title: "figure6 allostery"
output:
  pdf_document: default
  html_document: default
date: "2026-01-28"
---

```{r setup}
base_dir<-"/base_dir"
knitr::opts_knit$set(root.dir = base_dir)
setwd(base_dir)
```

```{r}
library(data.table)
library(ggplot2)
library(ggpubr)
library(readxl)
library(scales)
library(tidyr)
library(ggrepel)
```

#tables used
```{r}
mutations_distances_annotation_median_rSASA_Binding_merged<-fread("Data_pdzext/energies_pdzext/thirdOrder_mutations_distances_annotation_median_rSASA_Binding_merged.txt")
signif_muts_bysite_distances_annotation_median_rSASA_Binding_merged<-fread("Data_pdzext/energies_pdzext/thirdOrder_signif_muts_bysite_distances_annotation_median_rSASA_Binding_merged.txt")
```

#transform exponencial decay function to linear curve log(y)=log(a)+b⋅x
```{r}
# Function to fit an exponential decay curve using a linear transformation
fit_exponential_curve_linear_transformed<-function(xvector,yvector,tit,plotfig=TRUE,writepar=FALSE){
  # Convert to data.table
data <- data.table(x = xvector, y = yvector)
  # Perform linear regression on log-transformed y
fit <- lm(log(y) ~ x, data = data)

summary(fit)
# Extract coefficients
ln_a <- coef(fit)[1]  # Intercept = ln(a)
b <- coef(fit)[2]    #slope (b is negative)

# Convert ln(a) back to a
a <- exp(ln_a)

# Display estimated parameters
cat("Estimated a",dataset_value,":", a, "\n")
cat("Estimated b",dataset_value,":", b, "\n")

# Return estimated parameters as a list
  return(list(a = a, b = b))
}
```

#fitting of decay curve and plots Figure 6a
```{r}
all_ddg_table_abs_allo_noBI<-data.table()#create a data.table with the decay curve fitting results below
  
   name <- "mutations_distances_annotation_median_rSASA_Binding_merged"
    dt <- get(name)
    dataset_values<-c("1st_order_ddg","calculated_ddg_N","calculated_ddg_C","calculated_ddg_NC")
    
for (dataset_value in dataset_values) {
  subset_dt <- dt[dataset == dataset_value]
   subset_dt$abs_total_mean_kcal_mol <- abs(subset_dt$total_mean_kcal_mol)
median_abs<-median(subset_dt$abs_total_mean_kcal_mol)
mad_abs <- mad(subset_dt$abs_total_mean_kcal_mol, constant = 1, na.rm = TRUE)  # constant=1 to not rescale
threshold_abs_median <- median_abs + mad_abs 
  
  
  valid_rows <- !is.na(subset_dt$scHAmin_ligand) & !is.na(subset_dt$total_mean_kcal_mol) & subset_dt$binding_interfase==F #the binding interface is not use for the fitting
  
  
  xvector_starting=subset_dt[valid_rows]$scHAmin_ligand#minimun distance from the side chain heavy atom to the ligand
  yvector_starting=abs(subset_dt[valid_rows]$total_mean_kcal_mol)#absolute ddg values
    
  # Fit the exponential decay model
    exponential_curve_fitted<-fit_exponential_curve_linear_transformed(xvector=xvector_starting,yvector=yvector_starting,tit=dataset_value,plotfig=F,writepar=FALSE)
    
     # Extract coefficients
  a_fit <- exponential_curve_fitted$a
  b_fit <- exponential_curve_fitted$b

    #this model is a*e**bx
    subset_dt[,allo_predicted:=a_fit*exp(b_fit*scHAmin_ligand)]
    subset_dt[,allo_decay_residual:=abs(total_mean_kcal_mol)-allo_predicted]
    subset_dt[binding_interfase==T,allo_decay_residual:=NA]
    subset_dt[,label_id:=wt_Pos]
    
    # ----- Spearman ONLY for scHAmin_ligand > 5 -----
  spearman_df <- subset_dt[
    scHAmin_ligand > 5 &
    !is.na(scHAmin_ligand) &
    !is.na(abs_total_mean_kcal_mol)
  ]
    cat("\nDataset:", dataset_value, "- Spearman (scHAmin_ligand > 5)\n")
    print(cor.test(spearman_df$scHAmin_ligand,
                   spearman_df$abs_total_mean_kcal_mol,
                   method = "spearman"))
    
    
   # Plot results
      p<-ggplot(subset_dt, aes(x=scHAmin_ligand, y=abs(total_mean_kcal_mol)))+
        geom_point(data=subset_dt[scHAmin_ligand>5 & Pos != "312"], color="black",alpha=0.2,size=2)+
        geom_point(data=subset_dt[scHAmin_ligand<=5], alpha=0.3, color="#ff7f7f", size=2)+#binding interface residues in red
         geom_point(data=subset_dt[Pos=="312"], alpha=0.7, color="magenta", size=2)+#mutations for residue 312(change for checking any residue)
      
      geom_line(data=subset_dt, aes(x=scHAmin_ligand, y=allo_predicted), color="#9999ff", size=1)+#decay curve
      geom_hline(yintercept=threshold_abs_median, color = "#9999ff", linetype = "dashed")+#threshold cut off
      theme_classic()+
         theme(aspect.ratio=1)+
        # Spearman label (only scHAmin_ligand > 5)
    stat_cor(
      data = spearman_df,
      aes(x = scHAmin_ligand, y = abs_total_mean_kcal_mol),
      method = "spearman",
      p.accuracy = 1e-100,
  r.accuracy = 0.01,
      colour  = "black",
      size = 6
    ) +
     
      xlab("minimum distance to the ligand (A)")+
      ylab("abs(total_mean_kcal_mol)")+
      ggtitle(paste0(dataset_value))
      
      print(p)
      #ggsave(paste0("Data_pdzext/output_pdzext/Fig_6a_",dataset_value,".pdf"),width=8, height=5)
      
      all_ddg_table_abs_allo_noBI<-rbind(all_ddg_table_abs_allo_noBI, subset_dt)#create a data.table with the decay curve fitting results
}
```

#find allosteric hotspots
```{r}
#i have into account the z_score for significantly different from 0 total_mean_kcal_mol (fdr_stab_destab)
#i defined fdr_stab_destab<0.05
#I do fisher test for enrichment in significant mutations by residue

#####here we choose the residues enriched in significant mutations as in figure 3 and 4
 name1 <- "signif_muts_bysite_distances_annotation_median_rSASA_Binding_merged"
    dt1 <- get(name1)
    dataset_values<-c("1st_order_ddg", "calculated_ddg_N", "calculated_ddg_C", "calculated_ddg_NC")
    
for (dataset_value in dataset_values) {
  subset_dt1 <- dt1[dataset == dataset_value]
  
n_sign_ddg<-subset_dt1[signif_enrichment==TRUE]
assign(paste0("n_sign_ddg_",dataset_value),n_sign_ddg)

#####here we perform the test for allostery testing for the distance of residuals to the decay curve
name2<-  "all_ddg_table_abs_allo_noBI"
    dt2 <- get(name2)
    dataset_all<-c("1st_order_ddg", "calculated_ddg_N", "calculated_ddg_C", "calculated_ddg_NC")

  subset_dt2 <- dt2[dataset == dataset_value]
test_results <- subset_dt2[
  !is.na(allo_decay_residual),  # Ensure allo_decay_residual is not NA so we perform the test only in non-BindingInterface residues
  .(
    p_value_allost = if (.N > 1 && length(unique(allo_decay_residual)) > 1) {
      t.test(allo_decay_residual, mu = 0, alternative = "greater")$p.value
    } else {
      NA_real_  # Use numeric NA for consistency
    }
  ),
  by = .(Pos)
]

test_results[, p_adjusted_allost := p.adjust(p_value_allost, method = "fdr")] #calculates fdr of being above the decay curve as residue, for all the residue even those that are above the curve just because the curve for far away values is very low. for that i need to filter by n_sign_ddg that takes into account the ddg distribution t-test

assign(paste0("test_results_",dataset_value),test_results)

# Merge with `n_sign_ddg` to keep only hotspot candidates
fisher_results_significant <- merge(test_results, n_sign_ddg, by = "Pos", all.x = TRUE)
fisher_results_significant[,significant_allost_residue:=F] 
fisher_results_significant[p_adjusted_allost<0.1 & !is.na(wt_Pos), significant_allost_residue:=T]


assign(paste0("fisher_results_significant",dataset_value),fisher_results_significant)

fisher_significant_distributions<-fisher_results_significant[significant_allost_residue==T]

# Assign the new table dynamically
  assign(paste0("fisher_significant_distributions_", dataset_value), fisher_significant_distributions)
}
```

#modify "all_ddg_table_abs_allo_noBI" to add columns showing allosteric significance
```{r}
#add fdr_signif to all_ddg_table_abs_allo_noBI
name <- "all_ddg_table_abs_allo_noBI"
dt <- get(name)

for (dataset_value in unique(dt$dataset)) {
  # Get the corresponding Fisher test results
  fisher_results_name <- paste0("fisher_significant_distributions_", dataset_value)
  fisher_results <- get(fisher_results_name)

  # Merge fdr_signif with the original dataset
  dt[dataset == dataset_value, fdr_signif := fisher_results[.SD, on = "Pos", fdr_signif]]
  dt[dataset == dataset_value, p_adjusted_allost := fisher_results[.SD, on = "Pos", p_adjusted_allost]]
  dt[dataset == dataset_value, significant_allost_residue := fisher_results[.SD, on = "Pos", significant_allost_residue]]
  dt[is.na(significant_allost_residue), significant_allost_residue:=F]
  dt[dataset == dataset_value, significant_allost_mutation := significant_allost_residue==TRUE&significant_mutation==TRUE]
}

# Save the updated table back
assign(name, dt)
```

#plot Figure 6b: allosteric hotspots violin plots
```{r}
custom_colors <- c("1st_order_ddg"="blue","calculated_ddg_N"="#d783ffff","calculated_ddg_C"="#d4fb79ff",  "calculated_ddg_NC"="#d6d6d6ff")

all_ddg_table_abs_allo_noBI[, dataset := factor(dataset, 
    levels = c("1st_order_ddg", "calculated_ddg_N", "calculated_ddg_C", "calculated_ddg_NC"))]

ggplot(all_ddg_table_abs_allo_noBI[binding_interfase==F&dataset %in% c("1st_order_ddg", "calculated_ddg_N", "calculated_ddg_C", "calculated_ddg_NC")])+
  geom_hline(aes(yintercept = 0), color="grey")+
  geom_violin(aes(x=factor(Pos), y=allo_decay_residual, fill=dataset, color=significant_allost_residue),position = position_dodge(width = 0.9))+
  geom_dotplot(aes(x=factor(Pos), y=allo_decay_residual, fill=dataset,color=significant_allost_residue),binaxis = 'y', stackdir = 'center', dotsize = 0.5, alpha = 0.7,position = position_dodge(width = 0.9)) +  # Add dotplot for individual points)+
  theme_pubr()+
  scale_fill_manual(values=custom_colors)+
  scale_color_manual("significant_allost_residue==TRUE", values=c("grey80", "black"))+
 
  facet_wrap(~Pos, scales="free_x", nrow=5)+
  ggtitle("significant_allost_residue==TRUE")+
  
  theme(legend.position="right", axis.text.x = element_blank(), axis.ticks.x = element_blank())
ggsave("Data_pdzext/output_pdzext/Fig_6b.pdf",width=15, height=15)
```

#plot figure 6c
```{r}
  name <- "mutations_distances_annotation_median_rSASA_Binding_merged"
    dt <- get(name)
    
    subset_dt <- dt[dataset %in% c("1st_order_ddg", "calculated_ddg_N", "calculated_ddg_C", "calculated_ddg_NC")]
     subset_dt[, dataset := factor(dataset, 
    levels = c("1st_order_ddg", "calculated_ddg_N", "calculated_ddg_C", "calculated_ddg_NC"))]
     positions <- c("312", "318", "329", "330", "336", "337", "341", "347", "357", "373", "375")
     subset1_dt <- subset_dt[Pos %in% positions]
     subset1_dt[, Pos := factor(Pos, levels = positions)]
    
   subset1_dt$mutant_colored<-factor(subset1_dt$mutant_colored,levels=c(
  "<span style='color:orange'>S</span>",                                                                    "<span style='color:orange'>T</span>",
  "<span style='color:orange'>N</span>",
  "<span style='color:orange'>Q</span>",
  "<span style='color:red'>D</span>",
  "<span style='color:red'>E</span>",
  "<span style='color:blue'>K</span>",
  "<span style='color:blue'>R</span>",
  "<span style='color:blue'>H</span>",
  "<span style='color:violet'>G</span>",
  "<span style='color:violet'>P</span>",
  "<span style='color:violet'>C</span>",
  "<span style='color:pink'>M</span>",
  "<span style='color:pink'>A</span>",
  "<span style='color:pink'>L</span>",
  "<span style='color:pink'>I</span>",
  "<span style='color:pink'>V</span>",
  "<span style='color:pink'>F</span>",
  "<span style='color:pink'>Y</span>",
  "<span style='color:pink'>W</span>"))    

   subset1_dt[, wt_Pos := factor(Pos, levels = positions)]
  subset1_dt <- complete(
  subset1_dt,
  wt_Pos,
  dataset,
  mutant_colored,
  fill = list(total_mean_kcal_mol = NA)
)
  subset1_dt <- as.data.table(subset1_dt)
subset1_dt[, is_na := is.na(total_mean_kcal_mol)]

p<-ggplot(subset1_dt,aes(x=dataset,y=mutant_colored,fill=total_mean_kcal_mol))+
     geom_tile()+
  geom_text(
    data = subset1_dt[is_na == TRUE],
    aes(label = "-"),
    color = "black", size = 5, na.rm = TRUE
  ) +
  geom_point(
   data = subset1_dt[class %in% c("stabilizing", "destabilizing")],
   aes(x = dataset, y = mutant_colored),
  shape = 18, color = "black", size = 0.5
  ) +
    facet_wrap(~ wt_Pos, nrow = 1) +
  scale_fill_gradientn(colours = c("#898ac0","#f5f5f5","#e85d47"),na.value = "#f5f5f5",guide = "colorbar", values = scales::rescale(c(-1.3, 0, 1.3)),limits = c(-1.3, 1.3)) +
    coord_fixed(ratio = 1)+    
     theme_classic()+
    theme(
  axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
  axis.text.y = element_markdown(),
  legend.key.size = unit(0.5, "cm"),
  legend.text = element_text(size = 8),
  legend.title = element_text(size = 0))+
       
  ggtitle("Energy term heatmaps for multiple positions")
    
ggsave(paste0("Data_pdzext/output_pdzext/Fig_6c.pdf"), width=15, height=5)
  print(p)
```

#plot figure 6e
```{r}
dataset_values<-c("1st_order_ddg", "calculated_ddg_N", "calculated_ddg_C", "calculated_ddg_NC")
name <- "all_ddg_table_abs_allo_noBI"
dt <- get(name)

for (dataset_value in dataset_values) {
   subset_dt <- dt[dataset == dataset_value]
   subset_dt_unique <- subset_dt[, .SD[1], by = Pos]
   
   label_dt <- subset_dt_unique[significant_allost_residue == TRUE]
   label_11hotspots <- subset_dt_unique[Pos %in% c("312","318","329","330","336","337","341","347","357","373","375")]
   
   subset1_dt<-subset_dt_unique[median_energy<0]
  subset2_dt<-subset_dt_unique[median_energy>0]
  
  
  p<-ggplot(subset_dt_unique, aes(x=abs(median_energy),y=rSASA_1be9))+
    geom_point(shape = 21,color="black", fill=NA,size=3)+
    geom_point(data = label_11hotspots,
                shape = 21,color="black", fill="darkgrey",size=5)+
     geom_point(data = subset_dt_unique[significant_allost_residue == TRUE],
           shape = 21, color = "black", fill = "red", size = 5)+
     
    
    geom_text(data = label_11hotspots,
              aes(x = abs(median_energy), y = rSASA_1be9, label = Pos),
              color = "black", size = 8, vjust = -1) +
    geom_text(data = label_dt,
              aes(x = abs(median_energy), y = rSASA_1be9, label = Pos),
              color = "red", size = 8, vjust = -1) +
    
     scale_color_manual(values = custom_colors) +
    geom_hline(yintercept = 0.2, color = "blue", linetype = "dashed") +
    
    # # Correlation for all
    stat_cor(data = subset_dt_unique, aes(x = abs(median_energy), y = rSASA_1be9), method = "spearman", color = "black",size=10,label.x = 0, label.y = max(subset_dt_unique$rSASA_1be9, na.rm = TRUE)*0.9,
           hjust = 0) +
    
    theme_classic()+
     theme(aspect.ratio=1)+
  theme(axis.text.x = element_text(size = 15),  # Adjust size of x-axis text
            axis.text.y = element_text(size = 15),  # Adjust size of y-axis text
            axis.title.x = element_text(size = 15), # Adjust size of x-axis title
            axis.title.y = element_text(size = 15),
            legend.position = "none") +
    ggtitle(dataset_value)
  print(p)
  ggsave(paste0("Data_pdzext/output_pdzext/Fig_6e_",dataset_value,".pdf"), width=8, height=5)
}
```
#plot figure 6g: test Pos 341 for +C vs -C
```{r}
  name <- paste0("mutations_distances_annotation_median_rSASA_Binding_merged")
    dt <- get(name)
    subset_dt<- dt[Pos==341&dataset%in%c("1st_order_ddg","calculated_ddg_N", "calculated_ddg_C", "calculated_ddg_NC"),]
    subset_dt[, dataset := factor(dataset, levels = c("1st_order_ddg", "calculated_ddg_N", "calculated_ddg_C", "calculated_ddg_NC"))]
    
group_map <- list(
  Group1 = c("1st_order_ddg", "calculated_ddg_N"),
  Group2 = c("calculated_ddg_C", "calculated_ddg_NC")
)
subset_dt[, group := fifelse(dataset %in% group_map$Group1, "Group1", "Group2")]
# Aggregate to one value per group per id
agg_dt <- subset_dt[, .(mean_energy = mean(total_mean_kcal_mol, na.rm = TRUE)), by = .(id, group)]
    # Reshape for paired t-test
wide_dt <- dcast(agg_dt, id ~ group, value.var = "mean_energy")
t_res <- t.test(wide_dt$Group1, wide_dt$Group2, paired = TRUE)
print(t_res)

p_value_label <- paste0("Paired t-test p = ", signif(t_res$p.value, 3))
    
p <- ggplot(subset_dt, aes(x = dataset, y = total_mean_kcal_mol)) +
 
 geom_point(width = 0.2, alpha = 0.7, size = 5, color="#e85d47") +
   
   geom_boxplot(width = 0.2, outlier.shape = NA, fill = NA, color = "black", size = 0.5) +
  geom_line(aes(group = id), color = "black", alpha = 0.3) +
  
  theme_classic() +
   theme(aspect.ratio=1)+
  ggtitle(p_value_label)

  print(p)
    ggsave(paste0("Data_pdzext/output_pdzext/Fig_6g.pdf"), width=7, height=7)
```

#table to compare better allosteric hotspots with rSASA. only residues outside binding interface (not BI)
```{r}
#reduce to unique Pos the table containing the info i want(was for mutations)
name <- ("all_ddg_table_abs_allo_noBI")
dt <- get(name)
subset_dt <- unique(dt[, .(dataset,Pos, rSASA_1be9,binding_interfase, allo_hotspot = significant_allost_residue)])
subset_dt<-subset_dt[binding_interfase==FALSE]

#add a column structure_location for core/surface
subset_dt[, structure_location := fifelse(rSASA_1be9 < 0.2, "core", "surface")]

#add a column if conserved allo_hotspot
conserved_allo_hotspots<-c("318","329","330","336","341","347") #defined as allo_hotspot in at least 3 context
subset_dt[, conserved_allo_hotspot := as.character(Pos) %in% conserved_allo_hotspots]

#add a column if not conserved allo_hotspot
subset_dt[, not_conserved_allo_hotspot := allo_hotspot==TRUE &!(as.character(Pos) %in% conserved_allo_hotspots)]

assign("allo_hotspots_table", subset_dt)
```

#fisher test allosteric hotspots enriched in core or not? for all the libraries together. 
```{r}
name<-  "allo_hotspots_table"
dt<-get(name)

dt[, allo_hotspot := factor(allo_hotspot, levels = c("TRUE","FALSE"))]
dt[, structure_location := factor(structure_location, levels = c("core","surface"))]
#table like contingency matrix for fisher test
allo_hotspots_counts <- with(dt, table(allo_hotspot, structure_location))
print(allo_hotspots_counts)

fisher_hotspots_core <- fisher.test(allo_hotspots_counts)
print(fisher_hotspots_core)

#which are the 8 allo_hotspots in the surface?
print(surface_allo <- dt[allo_hotspot == TRUE & structure_location == "surface", .(Pos, dataset)])

#odds_signif
allo_hotspots_in_core_or<-fisher_hotspots_core$estimate
allo_hotspots_in_surface_or<-1/allo_hotspots_in_core_or

#p_value
allo_hotspots_in_core_p<-fisher_hotspots_core$p.value

plot_df <- data.frame(
  group    = c("Core", "Surface"),
  log2_or  = c(log2(allo_hotspots_in_core_or), log2(allo_hotspots_in_surface_or)),
  p_value = allo_hotspots_in_core_p
)

p_value_label <- paste0("one-sided Fisher’s exact test p = ", signif(fisher_hotspots_core$p.value, 3))

ggplot(plot_df, aes(x = group, y = log2_or)) +
  geom_col(width = 0.2, show.legend = FALSE) +
  
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = NULL, y = "log2(odds ratio)", title = "Core vs Surface enrichment for allo_hotspots",
       subtitle = p_value_label) +
  theme_classic(base_size = 12) +
  theme(plot.title.position = "plot")
#ggsave(paste0("Data_pdzext/output_pdzext/Extended_Data_Fig_4b.pdf"), width=8, height=5)

```


#violin plot for rSASA in hotspots  compared with not hotspots
```{r}
#custom_colors <- c("surface" = "#17A2B8", "core" = "#94da55", "binding_interface"="#ff7f7f")


  name <- "allo_hotspots_table"
    dt <- get(name)
   
  
  
  dt[, allo_hotspot := factor(
  allo_hotspot,
  levels = c("FALSE", "TRUE")
)]
  
  counts <- dt[, .N, by = allo_hotspot]
print(counts)
    
    comparisons_list <- list(
      c("FALSE", "TRUE")
    )

 y_max <- max(dt$rSASA_1be9, na.rm = TRUE)
y_lim <- y_max * 1.5  # or add a fixed buffer if you prefer: y_max + 1 

p <- ggplot(dt, aes(x = allo_hotspot, y = rSASA_1be9)) +
  geom_jitter(aes(color = allo_hotspot), width = 0.2, alpha = 0.7, size = 0.1) +
  geom_violin(aes(fill = allo_hotspot), trim = TRUE, alpha = 0.3, size = 0.1) +
 #scale_color_manual(values = custom_colors) +
  #scale_fill_manual(values = custom_colors) +
  stat_compare_means(method = "t.test", 
                           comparisons = comparisons_list,
                           label = "p.format",
                           size = 10)+
 scale_y_continuous(limits = c(min(dt$rSASA_1be9, na.rm = TRUE), y_lim)) +
  theme_classic() +
   theme(aspect.ratio=1)+
   geom_hline(yintercept = 0.2, color = "blue", linetype = "dashed") 
  
  #labs(
   # title = "two-sided unpaired Student’s t-tests; **** = p ≤ 0.0001"
    
  #) 
  

  print(p)
#ggsave(paste0("Data_pdzext/output_pdzext/Extended_Data_Fig_4c.pdf"), width=8, height=5)

```


#fisher test allosteric hotspots enriched in core or not? one test for each library:
```{r}
name <- "allo_hotspots_table"
dt <- get(name)


dt[, allo_hotspot := as.character(allo_hotspot) == "TRUE",levels = c("TRUE","FALSE")]

dt[, structure_location := factor(structure_location,
                                  levels = c("core","surface"))] 
# Build table per dataset
counts_by_dataset <- dt[, as.data.table(table(allo_hotspot, structure_location)), by = dataset]

# Reshape wide
counts_wide <- dcast(counts_by_dataset, dataset + allo_hotspot ~ structure_location, value.var = "N")
counts_wide[, allo_hotspot := factor(allo_hotspot, levels = c("TRUE","FALSE"))]
setorder(counts_wide, dataset, allo_hotspot)

print(counts_wide)


# 3) Per-dataset Fisher’s exact test (two-sided)
fisher_by_dataset <- dt[, {
  tab <- table(allo_hotspot, structure_location)  # rows: TRUE/FALSE; cols: core/surface
  fres <- fisher.test(tab)                        # two-sided
  data.table(
    odds_ratio   = as.numeric(fres$estimate),
    log2_or      = log2(as.numeric(fres$estimate)),
    p_two_sided  = fres$p.value
  )
}, by = dataset]

print(fisher_by_dataset)

fisher_by_dataset[,log2_or_surface:=-log2_or]



custom_colors <- c("1st_order_ddg"="blue","calculated_ddg_N"="#d783ffff","calculated_ddg_C"="#d4fb79ff",  "calculated_ddg_NC"="#d6d6d6ff")
plot_df <- rbind(
  fisher_by_dataset[, .(dataset, facet = "Core",    y =  log2_or, p_two_sided)],
  fisher_by_dataset[, .(dataset, facet = "Surface", y = -log2_or, p_two_sided)]
)

# Truncate ±Inf to a finite cutoff for plotting (and optionally label)
cutoff <- 5
plot_df[, y_plot := fifelse(is.infinite(y), ifelse(y > 0,  cutoff, -cutoff), y)]
plot_df[, trunc_label := fifelse(is.infinite(y), ifelse(y > 0, paste0("> ", cutoff), paste0("< -", cutoff)), NA_character_)]

plot_df[,dataset:=factor(dataset,levels = c("1st_order_ddg", "calculated_ddg_N", "calculated_ddg_C", "calculated_ddg_NC"))]
plot_df[, significant_p := p_two_sided < 0.05]



ggplot(plot_df, aes(x = dataset, y = y_plot, fill = dataset,color=significant_p)) +
   scale_fill_manual(values=custom_colors)+
  scale_color_manual("p<0.05", values=c("grey80", "black"))+
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  # label truncated bars
  geom_text(aes(label = trunc_label), vjust = -0.4, na.rm = TRUE, size = 3) +
  facet_wrap(~ facet, ncol = 2, scales = "fixed") +
  labs(
    x = NULL,
    y = "log2(OR)",
    title = "Core vs Surface enrichment among allo_hotspots by dataset",
    subtitle = paste0("Bars truncated at ±", cutoff)
  ) +
  theme_classic(base_size = 12) +
  theme(plot.title.position = "plot",
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

```




#fisher test more conserved allosteric hotspots enriched in core or not? for all the libraries together.
```{r}
name<-  "allo_hotspots_table"
dt<-get(name)

dt[, conserved_allo_hotspot := factor(conserved_allo_hotspot, levels = c("TRUE","FALSE"))]
dt[, structure_location := factor(structure_location, levels = c("core","surface"))]
#table like contingency matrix for fisher test
conserved_allo_hotspots_counts <- with(dt, table(conserved_allo_hotspot, structure_location))
print(conserved_allo_hotspots_counts)

fisher_conserved_hotspots_core <- fisher.test(conserved_allo_hotspots_counts)
print(fisher_conserved_hotspots_core)

#which are theconserved_ allo_hotspots in the surface?
print(surface_conserved_allo <- dt[conserved_allo_hotspot == TRUE & structure_location == "surface", .(Pos, dataset)])

#odds_signif
allo_hotspots_in_core_or<-fisher_hotspots_core$estimate
allo_hotspots_in_surface_or<-1/allo_hotspots_in_core_or

#p_value
allo_hotspots_in_core_p<-fisher_hotspots_core$p.value
```

#fisher to test if more conserved hot spots tend to be in the core more than not_conserved
```{r}
name="allo_hotspots_table"
dt<-get(name)

subset_dt<-dt[conserved_allo_hotspot%in%TRUE | not_conserved_allo_hotspot%in%TRUE]
subset_dt[, conserved := conserved_allo_hotspot%in%TRUE]

subset_dt[, conserved := factor(conserved, levels = c(TRUE,FALSE))]
subset_dt[, structure_location := factor(structure_location, levels = c("core","surface"))]

#table like contingency matrix for fisher test
conserved_core <- with(subset_dt, table(conserved, structure_location))
print(conserved_core)

fisher_conserved_core <- fisher.test(conserved_core)
print(fisher_conserved_core)

```

######################
###enrichment in sector residues
########################

```{r}
sector_positions <- c(329, 330, 336, 338, 341, 347, 353, 359, 362, 367, 375)
allo_hotspots_table[, sector := Pos %in% sector_positions]
allo_hotspots_table[, {
  tab <- table(allo_hotspot, sector)
  list(
    p_value = fisher.test(tab)$p.value,
    odds_ratio = fisher.test(tab)$estimate
  )
}, by = dataset]
```

